---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidyRename

<!-- badges: start -->
  [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
  <!-- badges: end -->




## Purpose

`{tidyRename}` was originally designed for automating the creation of submission compliant variable attributes for SAS v5 Transport files (i.e. <=8 characters for variable names, among other things). However, the same concepts can be expanded to CSR activities i.e., SDTM variables mapping process etc. to ensure uniformity among users and eliminates wasted time manually creating easily-readable variable names for high dimensional data sets. In fact, `{tidyRename}` can be useful outside of the pharma industry as well - it's a handy "renamer". Just give it a vector of strings, and it will tidy them up for you to meet your custom needs. See example use cases below!
    
<br>

    
## Origins

Project built off of this [contribution](https://github.com/atorus-research/xportr/pull/11/files) to [{xportr}](https://github.com/atorus-research/xportr) from the [{pharmaverse}](https://pharmaverse.org/)

<br>


## What are the checks?

If you interested in using `{tidyRename}` for submission compliance to regulatory bodies, here's what you can expect from it's output:


- Variable names must start with a letter (not an underscore), be comprised of only uppercase letters (A-Z), numerals (0-9) and be free of non-ASCII characters, symbols, and underscores.
- Allotted length for each column containing character (text) data
should be set to the maximum length of the variable used across all
data sets (≤ 200)
- Coerces variables to only numeric or character types
- Display format support for numeric float and date/time values
- Variables names are ≤ 8 characters.
- Variable labels are ≤ 200 characters. 
- Data set labels are ≤ 40 characters.
- Presence of non-ASCII characters in Variable Names, Labels or data set labels.

<br>


## Installation

You can install the development version of tidyRename from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("aclark02-arcus/tidyRename")
```


<br>

## Example use cases

### Renaming strings

This is a basic example which shows you how to solve a common problem with raw data variable names, manipulated as a character vector:

```{r example1}
library(tidyRename)

## basic example converting 
vars <- c("", "studyid", "STUDYID", "subject id", "1c. ENT", "1b. Eyes",
          "1d. Lungs", "1e. Heart", "year number", "1a. Skin_Desc")

# Default behavior
tidy_rename(vars)

```

The user can alter the final letter case of the output by specifying "upper", "lower", or "asis" to maintain the input string case:

```{r example2}
# maintain source case
tidy_rename(vars, letter_case = "asis") 
```

By default, SAS v5 transport files require variable names to be 8 or less characters, but `{tidyRename}` can accept custom character lengths for the final output:

```{r example3}
# Set Custom character length in output
tidy_rename(vars, char_len = 10)
```

Usually, you don't want a number to be the first character of your variable name. By default, those numbers will be relocated to the end of the string. However, if you want to keep them up front, that's possible! But we recommend you add a letter prefix as seen below:

```{r example4}
# Leave numerical prefix on left-hand side, but add a starting letter
tidy_rename(vars, relo_2_end = FALSE, letter_for_num_prefix = "q")
```

Sometimes you'll have some common variable names that you'll always want to change to a standard name. For that use case, just specify a "dictionary data.frame", as seen in the below example:

```{r example5}
# Add a dictionary and remove underscores - impacts var 4
tidy_rename(vars, sep = "",
dict_dat = data.frame(original_varname = "subject id", dict_varname = "subjid"))
```

### Renaming data frame variable attributes directly

These basic examples shows you how to use `{tidyRename}` to manipulated a data.frame object's variable attributes directly, returning the input data.frame:


```{r example6}
vars <- c("", "STUDYID", "studyid", "subject id", "1c. ENT", "1b. Eyes",
      "1d. Lungs", "1e. Heart", "year number", "1a. Skin_Desc")
adxx <- data.frame(matrix(0, ncol = 10, nrow = 3))
colnames(adxx) <- vars


tidy_varnames(adxx) # default

tidy_varnames(adxx, relo_2_end = FALSE, letter_for_num_prefix = "q") # prefix numbers on left-hand side

my_dictionary <- data.frame(original_varname = "subject id", dict_varname = "subjid")
tidy_varnames(adxx, dict_dat  = my_dictionary) # 'SUBJID' used

tidy_varnames(adxx, sep = "_") # permissible for legacy studies
```





<br>


<br>


<br>


<br>


